---
title: "ATSA Term Project"
author: "AE14B051: Abhishek Kelkar"
date: "30 November 2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
This chunck of code checks for the Granger causality using Sij statistic provided for the given dataset. But first we need to find the values a{ij}

```{r Ahat}  
Ahat <- function(Data,Order)
{
  # Read input arguments
  Z = Data
  P = Order
    # Size of the data
  Size = dim(Z)
  N = Size[1]
  M = Size[2]
  # The regressor matrix
  BigZ = {}
  BigZ1 = {}
  
  for (k in 1:P){
    BigZ = cbind(BigZ, Z[k:(N-P+k-1),])
  }
  
  # Re-arrange the regressor matrix
  for (j in 1:M){
    BigZ1 = cbind(BigZ1,BigZ[,(seq(j,M*P,M))])
  }
  # The Y matrix   
  Y = Z[(P+1):N,]
  
  # Regressors covariance matrix and its inverse
  SigmaR = (t(BigZ1)%*%BigZ1)/N
  invSigmaR = qr.solve(SigmaR)
  
  # Estimate VAR model coefficient matrices
  Ahat = (qr.solve(t(BigZ1)%*%BigZ1)%*%t(BigZ1))%*%Y
  
  Amatrix = array(0,dim = c(M,M,P))
  for (i in seq(1,P)){
    Amatrix[,,i] = Ahat[((i-1)*M+1):(i*M),1:M]
  }
  return(Amatrix)
}

```
The Vhat matrix function is provided before hand as follows
```{r vhat}
Vhat <- function(Data,Order)
{
  # Function to obtain V_ij matrix (P*P) from the estimate of covariance matrix (M^2P*M^2P) of
  # VAR model coefficients
  # Reference: Michael Eichler (2005) and Lutkepohl H (1993)
  # Usage: [Vmat] = Vhat(Data, Order)
  # Sudhakar Kathari, November 11, 2017.
  
  # Read input arguments
  Z = Data
  P = Order
  
  # Size of the data
  Size = dim(Z)
  N = Size[1]
  M = Size[2]
  
  # The regressor matrix
  
  BigZ = {}
  BigZ1 = {}
  
  for (k in 1:P){
    BigZ = cbind(BigZ, Z[k:(N-P+k-1),])
  }
  
  # Re-arrange the regressor matrix
  for (j in 1:M){
    BigZ1 = cbind(BigZ1,BigZ[,(seq(j,M*P,M))])
  }
  # The Y matrix   
  Y = Z[(P+1):N,]
  
  # Regressors covariance matrix and its inverse
  SigmaR = (t(BigZ1)%*%BigZ1)/N
  invSigmaR = qr.solve(SigmaR)
  
  # Estimate VAR model coefficient matrices
  Ahat = (qr.solve(t(BigZ1)%*%BigZ1)%*%t(BigZ1))%*%Y
  
  # Innovations covariance matrix
  SigmaE = (1/(N-P))*t((Y-BigZ1%*%Ahat))%*%(Y-BigZ1%*%Ahat)
  
  # Covariance matrix of VAR model coefficients
  SigmaA = kronecker(invSigmaR,SigmaE)
  
  
  # Estimation of V matrix
  Vmatrix = array(0,dim = c(P,P,M*M))
  for (i in seq(1,M*M)){
    Vmatrix[,,i] = SigmaA[((i-1)*P+1):(i*P),((i-1)*P+1):(i*P)]
  }
  return(Vmatrix)
}
```

Thus we can identify the Granger causality in section A
```{r Section A}
zk<-load("G:/IIT MADRAS/SEMESTER 7/CH5350 Appllied Time Series Analysis/Causality Analysis Project/proj2017.RData")
Vmatrix <- Vhat(Zk, 2)
Amatrix <- Ahat(Zk, 2)

M <- dim(Amatrix)[1]
P <- dim(Amatrix)[3]
s <- c()
N = nrow(Zk) #no. of observations in data
for (i in seq(1,M*M))
{
  aij <- c()
  for (j in seq(1,P))
  {
    aij <- append(aij, Amatrix[,,j][i])
  }
  
  sij <- N*matrix(aij, nrow = 1) %*% Vmatrix[,,i] %*% matrix(aij, ncol = 1)
  s <- append(s, sij)
}

Smatrix <- matrix(s, nrow = M)
Smatrix
```

The following code is for the section B to identify the direct pathway function for the given data-set Zk where N=2046 and M=3. The function below takes the A matrix from the VAR model which is of the order (MN \times MN) for all values of r and p is the order of the VAR model

```{r Section B}
Amatrix=Ahat(Zk,2)
Dim=dim(Amatrix)
sum<-dim(Amatrix)[1]
I=diag(sum)
w <- seq(0,pi,0.01)
PSI <- array(0, dim = c(Dim[1],Dim[2],length(w)))
  
for(i in w){
  Ar <- matrix(seq(0,0,length.out = Dim[1]^2), nrow = Dim[1])
  for(j in c(1:Dim[3])){
    Ar <- Ar + complex(1, cos(j*i), -sin(j*i)) * Amatrix[,,j]
    }
  Abarw=I-Ar
  
  hDij <- matrix(seq(0,0,length.out = Dim[1]^2), nrow = Dim[1])
  for(k in 1:dim(Abarw)[1]){
    for(l in 1:dim(Abarw)[1]){
      if(k==l){
        hDij[k,l] <- (det(Re(Abarw[-k,-l])))/det(Re(Abarw))
      }
      else{
        hDij[k,l] <- (-Abarw[k,l] * det(Re(Abarw[-k,-l])))/det(Re(Abarw))
      }
    }
  }
  
  PSIij <- matrix(seq(0,0,length.out = dim(Amatrix)[1]^2), nrow = dim(Amatrix)[1])
  for (m in c(1:dim(Abarw)[1]))
  {
    for (n in c(1:dim(Abarw)[1]))
    {
      PSIij[m,n] <- hDij[m,n] / sqrt(sum(hDij[,n]))
    }
  }
  
PSI[,,i] = PSIij
}

DPF <- function(coeff_matrix, w)
{
  amat <- coeff_matrix
  I <- diag(dim(amat)[1])
}
```


Frequency domain analysis for Granger Stability 

```{r Section C and D}
freq_domain_test <- function(data, order)
{
  data <- data
  order <- order
  N <- dim(data)[1]
  M <- dim(data)[2]
  DFT <- fft(data)
  
  R <- c(1:10)
  dpf_vector <- array(0, dim = c(M, M, 100, 10))
  psidist <- array(0, dim = c(M*M, 10, 100))
  zeta <- array(0, dim = c(M, M, 100))
  
  for (i in R)
  {
    ran <- runif(1,0,2*pi)
    mod_DFT <- DFT* complex(real = cos(ran), imaginary = sin(ran))
    
    surrogate <- fft(mod_DFT, inverse = TRUE)
    coeff_matrix <- Ahat(surrogate,order)
    
    dpf <- DPF(coeff_matrix, seq(0,2*pi, length.out = 100))
    dpf_vector[,,,i] <- dpf
    
  }
  for (i in 1:100){
      for (j in 1:M){
        for (k in 1:M){
          zeta[j,k,i] <- quantile(dpf_vector[j,k,i,], probs = 0.95)
        }
      }
    }
  return(zeta)
  
}
```

With N=2000 observations and a given generating process we now test our code

```{r Section E}
N=2000
I=diag(3)
A1=matrix(c(0.3,0,0,0.6,0.4,0,0,0.4,0.5),nrow=3,ncol=3)
A2=matrix(c(0.2,0,0,0,0.5,0,0,0.3,0.4),nrow=3,ncol=3)

#m1=VARMAsim(2000, arlags=c(1,2), phi=cbind(phi1,phi2), sigma=sig)
#d<- m1$series
#p <- vars(d, ic="SC")
```

**Section F**
When we run the frequency domain test function for Zk data-set from the given file proj2017.rdata, the following is obtained
* The modulus value of the DPF is 37.96343
 
```{r}
det(Re(PSIij*Conj(PSIij)))*det(Re(PSIij*Conj(PSIij)))
```

* The frequency domain analysis gives the value of the zeta function to be 1
* Since the modulus of the DPF is not greater than the zeta function, we can infer that z_j causes z_i
* On the other hand, the time domain analysis using the Sij matrix leads to the following
```{r}
qchisq(Smatrix,0.95)
```

*Thus we fail to reject the null hypothesis and hence we can claim that j is the cause of i
* Granger analysis results are consistent in time and frequency domain.

* The Granger causality analysis across realizations*
* Considering a physical view point, the cause and the effect should, ideally, remain same across multiple realizations and hence the Granger analysis should yield the same result.
* Granger causality measures can be highly sensitive to variations in the estimated model parameters, particularly in the frequency domain, leading to spurious peaks and valleys and even negative values.
* Causality measures are functions of the VAR system parameters and reflect a combination of dynamics from the different components of the system.
*Granger causality estimates are independent of the receiver dynamics, they reflect a combination of the transmitter and channel dynamics.
*Thus the answer can be expected to be invariant of the realizations. 

**Section 2**
The above functions are run for the data provided in Zk itself and the obtained results show that 
Zk[2] Granger causes Zk[1].